---
title: "p8105_hw5_sz3214"
output: github_document
date: "2024-11-09"

---

```{r setup, include=FALSE}
library(ggplot2)
library(broom)
library(readr)
library(dplyr)
library(purrr)
library(tidyr)
```


## Problem 1

**function**

```{r}
birthday = function(n) {
  birthdays = sample(1:365, n, replace = TRUE) 
  duplication = any(duplicated(birthdays))  
  return(duplication)
}

```

**run function on different group sizes**

```{r}
set.seed(42)  
n_simulation = 10000  
size = 2:50  
result = numeric(length(size))

for (i in seq_along(size)) {
  n = size[i]
  duplicate = replicate(n_simulation, birthday(n))  
  result[i] = mean(duplicate)  
}

data = data.frame(group_size = size, probability = result)
```

**draw the plot**

```{r}
ggplot(data, aes(x = group_size, y = probability)) +
  geom_line() +
  geom_point() +
  labs(title = "Probability of Shared Birthday in a Group",
       x = "Group Size",
       y = "Probability of Shared Birthday") +
  theme_minimal()
```

As the group size increases, the probability that at least two people in the group share a birthday rises quickly.


## problem 2

**t-test**

```{r}
n = 30  
sigma = 5 
alpha = 0.05 
num = 5000 
mu_value = c(0, 1, 2, 3, 4, 5, 6)

power = numeric(length(mu_value))
avg_estimates = numeric(length(mu_value))
avg_estimates_rejected = numeric(length(mu_value))

for (j in seq_along(mu_value)) {
  mu = mu_value[j]
  p_value = numeric(num)
  mu_hat = numeric(num)
  
  for (i in 1:num) {
    x = rnorm(n, mean = mu, sd = sigma)  
    t_test = t.test(x, mu = 0)  
    p_value[i] = t_test$p.value
    mu_hat[i] = t_test$estimate
  }
  
  power[j] = mean(p_value < alpha)
  avg_estimates[j] = mean(mu_hat)
  avg_estimates_rejected[j] = ifelse(sum(p_value < alpha) > 0, mean(mu_hat[p_value < alpha]), NA)
}

```

**plot mu vs power**

```{r}
data.frame(mu = mu_value, power = power) |>
  ggplot(aes(x = mu, y = power)) +
  geom_line() +
  geom_point() +
  labs(title = "Power vs. True Mean (mu)", x = expression(mu), y = "Power")
```

Effect size and power are positively correlated. 
The rate of increase in power slows as the effect size becomes very large.

**plot avg mu_hat vs mu**

```{r}
data.frame(mu = mu_value, avg_estimate = avg_estimates, avg_estimate_rejected = avg_estimates_rejected) |>
  ggplot(aes(x = mu)) +
  geom_line(aes(y = avg_estimate), color = "black") +
  geom_point(aes(y = avg_estimate), color = "black") +
  geom_line(aes(y = avg_estimate_rejected), color = "blue", na.rm = TRUE) +
  geom_point(aes(y = avg_estimate_rejected), color = "blue", na.rm = TRUE) +
  labs(title = "Average Estimate of mu_hat vs True Mean", x = expression(mu), y = "Average Estimate of mu_hat")
```

The average estimate of $\hat{\mu}$ for tests where the null hypothesis is rejected is approximately equal to the true value of $\mu$ because these tests provide stronger evidence of a real effect, making the estimates more accurate.


## problem 3


```{r}
homicide = read.csv("homicide-data.csv")
str(homicide)
```

The raw dataset has 52179 observations and 12 variables, including uid, reported_date, victim_last, victim_first, victim_race, victim_age, victim_sex, city, state, lat, lon and disposition.

**create new variable**   

```{r}
homicide_count = homicide |>
  mutate(state = ifelse(city == "Tulsa" & state == "AL", "OK", state)) |>
  mutate(city_state = paste(city, state, sep = ", ")) |>
  group_by(city_state) |>
  summarise(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest"))
  )

homicide_count
```

**Baltimore**

```{r}
homicide_count |>
  filter(city_state == "Baltimore, MD") |>
  summarise(prop_test = list(prop.test(unsolved_homicides, total_homicides))) |>
  mutate(output = map(prop_test, broom::tidy)) |>
  unnest(output) |>
  select(estimate, conf.low, conf.high)
```

**each city**

run prop.test

```{r}
each_city = homicide_count |>
  mutate(
    prop_test = purrr::map2(unsolved_homicides, total_homicides, ~prop.test(.x, .y) |> broom::tidy()),
    estimate = purrr::map(prop_test, ~.x$estimate),
    conf.low = purrr::map(prop_test, ~.x$conf.low),
    conf.high = purrr::map(prop_test, ~.x$conf.high)
  ) |>
  unnest(c(estimate, conf.low, conf.high)) |>
  select(city_state, estimate, conf.low, conf.high)

each_city
```

plot

```{r}
each_city |>
  ggplot(aes(x = reorder(city_state, estimate), y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  labs(title = "Proportion of Unsolved Homicides by City", x = "City", y = "Proportion of Unsolved Homicides") +
  theme_minimal()
```


